### 2020-09-29 paper/natastron/methods.tex
Done with the main body pretty much

For our sample, we used the asteroseismic ages obtained by \texttt{BASTA} \cite[BAyesian STellar Algorithm]{silvaaguirre+2015} in the Kages and LEGACY catalogues. These ages have been obtained by comparisons of measured oscillation properties to stellar models, accounting for an expanded range of metallicities. \texttt{BASTA} is thoroughly compared to four other seismic modelling techniques in \cite{silvaaguirre+2017}. While uncertainties found through \texttt{BASTA} are typically higher than for other techniques, only \texttt{BASTA} and \texttt{ASTFIT} \cite[Aarhus STellar Evolution Code]{christensen-dalsgaard2008} recover the radius, mass and age of the Sun, when applied to solar data. Although the uncertainties on \texttt{ASTFIT} ages are overall lower, they are not reported in \cite{silvaaguirre+2015} for the Kages sample. In order to maintain an internally consistent stellar age sample, we used age results from \texttt{BASTA} for both the Kages and LEGACY samples.

For our stellar masses we used asteroseismic model masses obtained by \texttt{BASTA} reported in Kages and LEGACY, in order to maintain internal consistency with the age measurements. We note that age and mass posteriors from \texttt{BASTA} are correlated, but chose not to account for the unpublished correlations in this work.
As described in the catalogue papers, for Kages stars atmospheric properties (\teff and \feh) were measured through high-resolution spectroscopy reported in \cite{huber+2013a}. For LEGACY stars, atmospheric properties were similarly taken from \cite{buchhave+latham2015} for most stars in the catalogue, and complemented by other values from the literature for the remaining stars \cite[see Table 3]{silvaaguirre+2017}.

\begin{figure*}
	\centering
	\includegraphics[width=.99\textwidth]{Images/modelfit.pdf}
	\caption{A power spectrum constructed from four years of \kepler observations of 16 Cyg A (KIC 12069424). Plotted over the top is the model resulting from the fit to the data described in this work. The model implements both the mode frequencies, seen on the right hand side of the plot, and the convective background, the effects of which are seen on the left. Low frequencies have been cropped out for clarity. \textit{Inset}: A zoom in on a radial (right) and quadrupole (left) ($\ell = 0, 2$) pair of modes. The quadrupole mode is split into five components by the star's rotation. Due to the star's inclination angle with respect to us, two out of five peaks are more distinct. The height and spacing of the mode components is a function of the star's rotational splitting ($0.56\, \mu\rm{Hz}$, equivalent to $P_{\rm rot} = 20.5\, \rm{days}$) and angle of inclination ($45^\circ$).}
	\label{fig:modelfit}
\end{figure*}

\section{Gyrochronology Methods}\label{s:gyro}
\subsection{Stellar Models}\label{ssec:models}
The braking models used in this work have several parameters that inform the rotational evolution. These are: a normalization factor to reproduce the solar rotation ($f_k$); a disk locking timescale ($T_{\rm disk}$) and period ($P_{\rm disk}$) which together regulate the stellar angular velocity during the pre-main sequence; the critical angular velocity that marks the transition from saturated (rapidly rotating) to unsaturated (slowly rotating) regimes ($\omega_{\rm crit}$), and the critical Rossby number, above which stars conserve angular momentum ($Ro_{\rm crit}$), mentioned above. $T_{\rm disk}$, $P_{\rm disk}$ and $\omega_{\rm crit}$ are calibrated to match the behavior in young open clusters, but have little impact on the rotational evolution beyond $\approx 1\, \rm{Gyr}$ in solar-mass stars. Both $f_k$ and $Ro_{\rm crit}$ affect the late-time evolution. Both models adopt $\omega_{\rm crit} = 3.4  \times 10^{-5}\, \rm{s}^{-1}$, $P_{\rm disk} = 8.1$, $T_{\rm disk} = 0.28$ and $f_k = 6.6$. In the weakened magnetic braking model, $Ro_{\rm crit} = 1.97$. For further details, see \cite{vansaders+pinsonneault2013} and \cite{vansaders+2016, vansaders+2019}.

Instead of comparing asteroseismic measurements of age and rotation directly to rotational models of individual stars, we instead compared them to synthetic populations based on rotational models, similar to those presented in \cite{vansaders+2019}, with some adaptations. To construct a synthetic population of rotating stars, \cite{vansaders+2019} applied the rotational braking laws described above to a galactic population model. Their approach used a TRILEGAL \cite{girardi+2012} Milky-Way simulation of the \kepler field, using the standard population values intended for this purpose from \cite{girardi+2015}. Stars from the TRILEGAL simulations were then drawn randomly in bins until they matched observations of stellar parameters of the \kepler field \cite{mathur+2017}. This aimed to replicate the \kepler selection effect, by ensuring the TRILEGAL population properties matched those observed in the \kepler field.

We used the same approach here, with two changes. First, we updated the distribution of stellar parameters matched to the TRILEGAL populations. Instead of \cite{mathur+2017}, we used the \cite{berger+2020} temperatures and luminosities, and 2MASS $K$-band magnitudes as our stellar parameter distributions for the \kepler field, which are more accurate following the use of \gaia DR2 parallaxes \cite{gaiacollaboration+2018}. The TRILEGAL simulation was matched to the \cite{berger+2020} sample using a nearest-neighbours approach (as opposed to binning), based on the density of stars on the HR diagram.

Second, we made further changes to account for possible binarity in the the \cite{berger+2020} sample. If the first step is performed blindly, blended binaries in the sample cause an overestimation of the number of old stars. In order to overcome this, we:

\begin{enumerate}
	\item blended the TRILEGAL stars with binary companions drawn from a flat mass-ratio distribution, using the binary fraction of \cite{raghavan+2010},
	\item recalculated the `observed' luminosities and magnitudes assuming that each binary pair was blended, and
	\item shifted these stars' temperatures following the $g$-$K$ relation presented in \cite{berger+2020}.
\end{enumerate}

@@

Our asteroseismic sample of stars with short cadence observations are subject to additional selection functions not included in the creation of the model populations above. We did \textit{not} explicitly account for these asteroseismic selection functions in our model, by design. Both the standard and WMB models contain stars with the same fundamental parameters (mass, radius, effective temperature, metallicity) but a different period based on the choice of rotational evolution prescription. Applying an asteroseismic selection function that depends on these fundamental parameters \cite{chaplin+2011} would apply an identical prior to both models, therefore providing no net effect on our posterior distribution. Additionally, we expected any seismic selection function to be relatively flat (and therefore uninformative) on a star-by-star scale, on which we run our model analysis (as opposed to a population scale, see below).

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{pgm_models.pdf}
	\caption{A probabilistic graphical model (PGM) represented algebraically in Equation \ref{eq:modelll}. The shaded circle indicates observed data, and solid black points represent other fixed information, such as the KDEs and observational uncertainties. The remaining circles represent parameters. The underline indicates that the symbol represents a set of parameters or data. Here, $\kappa_{\rm{s}}$ and $\kappa_{\rm WMB}$ represent the KDEs of standard and WMB model populations respectively. $P_{\rm{s}}$ is the mixture model weighting factor. The latent parameters $\underline{\theta}$, our observations $\underline{\mathcal{D}}$ and their uncertainties $\underline{\sigma_{\mathcal{D}}}$ include temperature (\teff), mass ($M$), log-age ($\ln(t)$), metallicity (\feh) and log-rotation ($\ln(P)$). This model is \textit{hierarchical}, as all the latent parameters are drawn from the common probability distribution set by $P_{\rm{s}}$ and described in Equation \ref{eq:mixturell}.}
	\label{fig:pgm}
\end{figure}

@@

Typically, this model would evaluate all stars in our sample against the stellar models simultaneously for a single posterior estimate of $P_{\rm s}$. At 95 stars, in 5 parameter spaces, this totals 476 free parameters to marginalise over. This is not an issue for Hamiltonian Monte Carlo \cite[HMC]{betancourt+girolami2013}, however the use of KDE functions, over which a probabilistic gradient can not be measured, reduces HMCs effectiveness. Alternative Markov Chain Monte Carlo techniques \cite[MCMC]{foreman-mackey+2013} can more efficiently sample the KDE functions, but can not treat the large number of hierarchical parameters. To overcome this, we fit our model to each star to obtain an independent individual posterior distribution for $P_{\rm s}$, and multiplied these post hoc to obtain a combined posterior. This comes with the benefit of easily allowing us to calculate the combined posterior for different stellar classifications (see Section \ref{s:class}), at the expense of the ability to marginalise for a single value of $P_{\rm s}$ directly.

\subsubsection{Fitting procedure}
The parameter space of the stellar models was reduced before calculating the KDEs. These cuts were made in $M$, $\teff$, $\ln(t)$ and \feh, removing any stars in the models that fell more than $3 \times \sigma_{\mathcal{D}}$ outside the observations. Our observables $M$, $t$ and $P$ have asymmetric uncertainties from the Bayesian asteroseismic analysis. We used the larger of the reported uncertainties on each parameter as $\sigma_{\mathcal{D}}$ in each parameter space.

@@

We fit our model Equation \ref{eq:modelll} using \texttt{emcee} \cite{foreman-mackey+2013}, using 32 walkers for a total of 7500 samples per walker, of which the first 2500 were discarded as a burn-in.

@@

\section{Verifying asteroseismic results}
\subsection{Priors on rotational parameters}
In our Bayesian analysis, we have placed weakly informative priors on our sampled rotational parameters, $\nu_{\rm s}\sin(i)$ and $\cos(i)$ (see Section \ref{s:seismo}). The prior is especially important for the angle of inclination, which is hardest to infer from the data. We are able to validate the robustness of our asteroseismic results by confirming that their posterior distributions are data-dominated, and not prior-dominated. We can do so by comparing the $68\%$ credible regions of the posterior estimates of $\nu_{\rm s}\sin(i)$ and $i$ against the $68\%$ credible regions of their priors.

A comparison between prior and posterior is shown for 94 stars in $\nu_{\rm s}\sin(i)$, $i$ and $P_{\rm rot}$ in Figure \ref{fig:priors}, arranged by age. In the Figure, results with means and error bars that closely resemble the prior distribution can be interpreted as prior-dominated (i.e. poorly informed by the data). The projected splitting, $\nu_{\rm s}\sin(i)$, is overall well constrained, with only one star appearing prior dominated. This is expected, as the projected splitting is what we observe on the star before decoupling inclination and rotation. The angle of inclination $i$, sampled as $\cos(i)$, more closely follows the prior distribution in most cases. Combining the two, the rotation period $P_{\rm rot}$ has no stars directly corresponding to the effective prior on period, and globally follows a trend with increasing age. The three outliers with fast rotation at late ages (KICs 6603624, 8760414 and 8938364) are discussed in more detail in Section \ref{ssec:litcomp}.

The rotation rates as presented in this work are a product of our Bayesian sampling of both projected splitting and angle of inclination. As seen in Figure \ref{fig:priors} there are instances where $i$ or $\nu_{\rm s}\sin(i)$ closely resemble the prior (and are therefore prior-dominated). There are no cases of this when looking at the resulting period measurements, as they will have been informed by at least one strongly data-driven parameter (judging from Figure \ref{fig:priors}, commonly $\nu_{\rm s}\sin(i)$). From this, we conclude that our ensemble of asteroseismic rotation is not strongly dominated by the priors imposed on projected splitting and inclination in our Bayesian analysis.

@@

Conversely, KICs 6603624, 8760414 and 8938364 have extremely slow rotation periods in LEGACY, but extremely \textit{fast} ($<3\, \rm {days}$) in this work. KICs 8760414 and 8938364 are excluded from the gyrochronology analysis below based on checks for $\hat{R}$ and the number of effective samples. Both stars have ages greater than $10\, \rm Gyr$, making their fast rotation rates highly unlikely under any model of rotational evolution. The posterior estimate of $P$ for KIC 6603624 is well-defined in our analysis, but with an age of $7.8\, \rm Gyr$, its measured rotation of $1.2\, \rm days$ is also highly unlikely under any model of rotational evolution. The LEGACY estimate of rotation is similarly extreme at $378\, \rm days$. These three stars have the lowest inclination angles in our sample ($< 10^\circ$), at which point the power in the split components of the seismic modes is so low that it becomes difficult to probe the measure of splitting\footnote{The split components for these stars will have a height roughly 3\% of the central mode. For comparison, for the next lowest inclined star at $17^\circ$, this rises to $10\%$.}. In cases such as these with lower signal-to-noise, spurious peaks may be interpreted as split components.\\

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{Images/litcomp_alt2.pdf}
	\caption{Comparisons between posterior estimates of rotational parameters from this work, LEGACY and Kages \cite[private communication]{davies+2016, lund+2017}. Fractional differences are plotted against stellar rotation obtained in this work. The $\Delta$ indicates the fractional difference betwen thsi work and the literature (i.e. stars above the zero-line have higher values in this work). The right hand panels the distribution of the fractional differences around the zero line. The colour legend is consistent throughout all panels. The x-axis units on the right hand panels are equivalent to the y-axis of the left hand panels. 10 stars have been omitted from this plot and are discussed in more detail in the text: KICs 5094751, 6196457, 8349582, 8494142, 8554498, 105114430 and 11133306 all have extremely low rotation periods in Kages, with high uncertainties. Conversely, KICs 6603624, 8760414 and 8938364 have extremely high rotation periods in LEGACY with low uncertainties. In cases where stars had asymmetric error bars, the larger of the two was used when propagating uncertainty for the purposes of this figure.}
	\label{fig:legacykages}
\end{figure*}

We also compared our asteroseismic estimates of stellar rotation with similar studies in the literature, shown in Figure \ref{fig:literaturecomp}. These included: \cite{davies+2015}, a study of the binary solar analogues 16 Cyg A \& B; \cite{nielsen+2015}, which our catalogue shares 5 stars with; and \cite{benomar+2018}, an asteroseismic study of differential rotation with which our catalogue shares 40 targets. For the latter, we used their reported splitting value $a_1$, which is equivalent to $\nu_{\rm s}$.

@@

It is of note that 16 Cyg A, as reported by \cite{davies+2015} was found to be rotating slightly faster in our analysis (deviating within $2\sigma$), despite the fit being performed on the same data. We found an inclination angle that is slightly lower than \cite{davies+2015} for 16 Cyg A but a similar projected splitting, which would explain finding a lower value of rotation.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.49\textwidth]{Images/seis_comparison_rot_alt2.pdf}
	\caption{Fractional differences between posterior estimates of asteroseismic rotation period from this work. Literature sources are: \cite{davies+2015} (16 Cyg A \& B), \cite{nielsen+2015} (5 stars) and \cite{benomar+2018} (40 stars). We used the reported parameter $a_{1}$ from \cite{benomar+2018}, which represents the rotational splitting in the case of uniform latitudinal rotation in their model. The dashed line represents the median of the sample shown, with the dotted lines representing the $15.9^{\rm{th}}$ and $84.1^{\rm{st}}$ percentiles. In cases where stars had asymmetric error bars, the larger of the two was used when propagating uncertainty for the purposes of this figure.}
	\label{fig:literaturecomp}
\end{figure}

\subsection{Internal vs external rotation}
Different methods of rotation measurement probe different regions of stars. Asteroseismology of main sequence stars probes internal rotation in the near surface layers, where the observed p modes are most sensitive \cite{lund+2014}. Measurements of star-spot modulation instead probe the rotation rates of star spots on the surface. A distinct difference between rotation rates obtained through these different techniques would hold information about differential rotation (both latitudinal and radial) of near-surface layers, such as those we observe in the Sun \cite{beck2000}.

\cite{nielsen+2015} performed a comparative analysis of 5 Sun-like stars observed with \kepler for which both star-spot modulation and asteroseismic rotation could be measured. They found no statistically significant difference between the two techniques.

Similarly, \cite{benomar+2015} performed an analysis on a larger sample of 22 stars. In this work they not only considered rotation rates from spots, but also spectroscopic measures of the projected surface rotation $\textrm{v}\sin(i)$, which they found to be more reliable. Comparisons with asteroseismic observations again showed no significant radial differential rotation.

With our expanded sample of asteroseismic rotation we can perform a similar analysis, to both validate our sample and probe radial differential rotation. Figure \ref{fig:vsinilit} shows a comparison between spectroscopic $\textrm{v}\sin(i)$ measurements as listed in LEGACY and Kages (left) and \cite{benomar+2015} (right). In these cases the asteroseismic $\textrm{v}\sin(i)$ has been calculated using our measure of asteroseismic $\nu_s\sin(i)$ and the known asteroseismic radii. Three stars (KICs 6603624, 8760414 and 8938364) have been excluded from this figure due to strong disagreements of measured rotation rates with the literature (see Section \ref{ssec:litcomp} above).

For the LEGACY and Kages sample, we find no strong deviation from the 1:1 line except at very low velocities, which is likely due to biases inherent to spectroscopic line broadening measurements \cite{doyle+2014, tayar+2015}. For the \cite{benomar+2015} sample stars lie a lot closer to the 1:1 line.
% , apart from one clear outlier. While we find no statistically significant difference between these observations, more in-depth studies of stars furthest away from the bisector in this sample may prove valuable in future work.

Overall, there appears to be a global offset where spectroscopic measurements of projected rotation appear faster than asteroseismic measures. Based on the LEGACY and Kages $v\sin(i)$ values, this offset is roughly $18\%$ (i.e. spectroscopic projected rotation rates are higher than asteroseismic rates). This offset is much smaller ($\sim5\%$) for the \cite{benomar+2015} sample, albeit for far fewer stars. These offsets are within the typical disagreement between spectroscopic methods, based on comparisons of projected rotation measurements for red giant stars \cite[see Figure 2]{tayar+2015}, especially at $< 5\, \rm{km\,s^{-1}}$.\\

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{Images/vsini-comparison.pdf}
	\caption{Comparisons between asteroseismic and spectroscopic measures of projected surface rotation, $\textrm{v}\sin(i)$. All asteroseismic (x-axis) values are from this work, all spectroscopic (y-axis) values are from the literature. \textit{Left}: comparisons to 81 stars values reported in LEGACY and Kages. \textit{Right}: comparisons to 16 stars observed by \cite{benomar+2015}. Asteroseismic values are transformed from projected splitting ($\nu_s\sin(i)$) using the asteroseismic radius measurements presented in LEGACY and Kages. The solid lines indicate the 1:1 line, while the dash-dotted lines represent the 2:1 and 1:2 lines.}
	\label{fig:vsinilit}
\end{figure*}

We also repeated the \cite{nielsen+2015} comparison of asteroseismic rotation rates to surface rotation rates, using surface rotation measurements from \cite{nielsen+2013} and \cite{garcia+2014}. Figure \ref{fig:protlit} shows comparisons for 48 stars with rotation rates from both techniques, including 4 from the original \cite{nielsen+2015} sample. Rotation from spot-modulation are subject to measuring multiples of the true rotation rate, unlike an asteroseismic analysis \cite{mcquillan+2014}, and therefore some stars may appear at the 2:1 and 1:2 lines on the Figure.

We repeated the analysis in \cite{nielsen+2015} by fitting a line of the form $P_{\rm surf} = m \times P_{\rm seis}$. For the purposes of this fit we used the larger of the asymmetrical uncertainties on the seismic rotation from this work. In order to avoid biasing the fit due to outliers, we only included stars below the 1.8:1 line.

Our fit found a value of $m = 0.96\pm0.03$, showing a close agreement ($< 2\sigma$) between asteroseismic and surface rotation rates on a population level. Of the 40 stars that were part of this analysis, 21 had a median value of $P_{\rm s} < 0.5$ in the gyrochronology analysis. We repeated the model fit for these stars only, and found a value of $m = 0.96 \pm 0.04$. As we find no statistically significant deviation, we conclude that our asteroseismic ensemble is in agreement with surface rotation measurements of these stars, and can therefore be used to draw inferences about gyrochronology.

The overall agreement within the ensemble between the photometric measures of surface rotation and the asteroseismic rotation rates is in line with previous studies of Sun-like stars \cite{gizon+2013, chaplin+2013, nielsen+2015, benomar+2015}, where different measures of rotation appear indistinguishable.
Individual deviations from this agreement, as appears to be the case for a number of stars that lie distinctly off the bisector, may warrant a more in-depth analysis, which we leave to future work.

@@

\section{Verifying consequences for gyrochronology}
\subsection{Rotational models for different stellar types}
Of the 73 stars for which posterior probability distributions of $P_{\rm s}$ were obtained, 4 were potential sub-giants, 22 were `hot' stars, and 47 were main sequence stars. Different stellar types should hold different amounts of diagnostic information about the standard or WMB models. Our `hot' stars should hold little to no information, as they lie above the Kraft break, where traditional gyrochronology relations used to build the stellar models break down (specifically, they do not lose angular momentum with time). Similarly, the rotational evolution of sub-giant stars is dominated by envelope expansion, complicating the interpretation of rotation.

>>>>

For our asteroseismic ages, we used the ages obtained by \texttt{BASTA} \cite[BAyesian STellar Algorithm]{silvaaguirre+2015} in the Kages and LEGACY catalogues. These ages have been obtained by comparisons of measured oscillation properties to stellar models, accounting for an expanded range of metallicities. \texttt{BASTA} is thoroughly compared to four other seismic modelling techniques in \cite{silvaaguirre+2017}. While uncertainties found through \texttt{BASTA} are typically higher than for other techniques, only \texttt{BASTA} and \texttt{ASTFIT} \cite[Aarhus STellar Evolution Code]{christensen-dalsgaard2008} recover the radius, mass and age of the Sun, when applied to solar data. Although the uncertainties on \texttt{ASTFIT} ages are overall lower, they are not published for the Kages sample. In order to maintain an internally consistent stellar age sample, we used age results from \texttt{BASTA} for both the Kages and LEGACY samples.

For our stellar masses we used asteroseismic model masses obtained by \texttt{BASTA} reported in Kages and LEGACY, in order to maintain internal consistency with the age measurements. We note that age and mass posteriors from \texttt{BASTA} are correlated, but chose not to account for the unpublished correlations in this work.

As described in the catalogue papers, for Kages stars atmospheric properties (\teff and \feh) were measured through high-resolution spectroscopy reported in \cite{huber+2013a}. For LEGACY stars, atmospheric properties were similarly taken from \cite{buchhave+latham2015} for most stars in the catalogue, and complemented by other values from the literature for the remaining stars \cite[see Table 3]{silvaaguirre+2017}.

Other asteroseismic properties used in this study, such as those used as first guesses on the free parameters in our asteroseismic model, are taken from the Kages and LEGACY catalogues \cite{astropycollaboration+2013, astropycollaboration+2018, ginsburg+2019}. In cases where both catalogues contained the same target, we used the stellar parameters reported in LEGACY. This is also the case for the masses and ages described above.

\subsection{Mode Frequency Fitting}

Todo: water down the contents of the methods here.

\section{Gyrochronology Methods}\label{s:gyro}
\subsection{Stellar Models}\label{ssec:models}
The braking models used in this work have several parameters that inform the rotational evolution. These are: a normalization factor to reproduce the solar rotation ($f_k$); a disk locking timescale ($T_{\rm disk}$) and period ($P_{\rm disk}$) which together regulate the stellar angular velocity during the pre-main sequence; the critical angular velocity that marks the transition from saturated (rapidly rotating) to unsaturated (slowly rotating) regimes ($\omega_{\rm crit}$), and the critical Rossby number, above which stars conserve angular momentum ($Ro_{\rm crit}$), mentioned above. $T_{\rm disk}$, $P_{\rm disk}$ and $\omega_{\rm crit}$ are calibrated to match the behavior in young open clusters, but have little impact on the rotational evolution beyond $\approx 1\, \rm{Gyr}$ in solar-mass stars. Both $f_k$ and $Ro_{\rm crit}$ affect the late-time evolution. Both models adopt $\omega_{\rm crit} = 3.4  \times 10^{-5}\, \rm{s}^{-1}$, $P_{\rm disk} = 8.1$, $T_{\rm disk} = 0.28$ and $f_k = 6.6$. In the weakened magnetic braking model, $Ro_{\rm crit} = 1.97$. For further details, see \cite{vansaders+pinsonneault2013} and \cite{vansaders+2016, vansaders+2019} for the standard and weakened magnetic braking cases respectively.

To construct a synthetic population of rotating stars, \cite{vansaders+2019} applied the rotational braking laws described above to a galactic population model. Their approach used a TRILEGAL \cite{girardi+2012} Milky-Way simulation of the \kepler field, using the standard population values intended for this purpose from \cite{girardi+2015}. Stars from the TRILEGAL simulations were then drawn randomly in bins until they matched observations of stellar parameters of the \kepler field \cite{mathur+2017}. This aimed to replicate the \kepler selection effect, by ensuring the TRILEGAL population properties matched those observed in the \kepler field.

To construct a synthetic population of rotating stars, we expanded upon previous forward-models of the \textit{Kepler} field with the purpose of studying weakened magnetic braking \cite{vansaders+2019}, with some important improvements. We started with a TRILEGAL \cite{girardi+2012} Milky-Way simulation of the \kepler field, using the simulation's standard population values intended for this purpose. In order to replicate the \kepler selection effect, the TRILEGAL simulation was matched to the largest current catalogue of temperatures, luminosities and 2MASS $K$-band magnitudes of stars in the \kepler field \cite{berger+2020}. This was done using a nearest-neighbours approach, based on the density of stars on the HR diagram. This ensured that the TRILEGAL population matched the density of stars actually observed by \kepler, replicating its selection effects.

We made further changes to account for possible binarity in the the matching sample \cite{berger+2020}. If the first step is performed blindly, blended binaries in the sample cause an overestimation of the number of old stars. In order to overcome this, we:

\begin{enumerate}
	\item blended the TRILEGAL stars with binary companions drawn from a flat mass-ratio distribution, using a known binary fraction \cite{raghavan+2010},
	\item recalculated the `observed' luminosities and magnitudes assuming that each binary pair was blended, and
	\item shifted these stars' temperatures following the $g$-$K$ relation presented in \cite{berger+2020}.
\end{enumerate}

@@

Our asteroseismic sample of stars with short cadence observations are subject to additional selection functions not included in the creation of the model populations above. We did \textit{not} explicitly account for these asteroseismic selection functions in our model, by design. Both the standard and WMB models contain stars with the same fundamental parameters (mass, radius, effective temperature, metallicity) but a different period based on the choice of rotational evolution prescription. Applying an asteroseismic selection function that depends on these fundamental parameters would apply an identical prior to both models, therefore providing no net effect on our posterior distribution \cite{chaplin+2011}.  Additionally, we expected any seismic selection function to be relatively flat (and therefore uninformative) on a star-by-star scale, on which we run our model analysis.

@@

@@

Typically, this model would evaluate all stars in our sample against the stellar models simultaneously for a single posterior estimate of $P_{\rm s}$. At 95 stars, in 5 parameter spaces, this totals 476 free parameters to marginalise over. This is not an issue for Hamiltonian Monte Carlo \cite[HMC]{betancourt+girolami2013}, however the use of KDE functions, over which a probabilistic gradient can not be measured, reduces HMCs effectiveness. Alternative Markov Chain Monte Carlo techniques \cite[MCMC]{foreman-mackey+2013} can more efficiently sample the KDE functions, but can not treat the large number of hierarchical parameters. To overcome this, we fit our model to each star to obtain an independent individual posterior distribution for $P_{\rm s}$, and multiplied these post hoc to obtain a combined posterior. This comes with the benefit of easily allowing us to calculate the combined posterior for different stellar classifications, at the expense of the ability to marginalise for a single value of $P_{\rm s}$ directly.

\subsubsection{Fitting procedure}
The parameter space of the stellar models populations were reduced before calculating the KDEs. These cuts were made in $M$, $\teff$, $\ln(t)$ and \feh, removing any stars in the models that fell more than $3 \times \sigma_{\mathcal{D}}$ outside the observations. Our observables $M$, $t$ and $P$ have asymmetric uncertainties from the Bayesian asteroseismic analysis. We used the larger of the reported uncertainties on each parameter as $\sigma_{\mathcal{D}}$ in each parameter space.

@@

We fit our model Equation \ref{eq:modelll} using \texttt{emcee} \cite{foreman-mackey+2013, foreman-mackey2016}, using 32 walkers for a total of 7500 samples per walker, of which the first 2500 were discarded as a burn-in.

@@

\section{Verifying asteroseismic results}
\subsection{Priors on rotational parameters}
In our Bayesian analysis, we have placed weakly informative priors on our sampled rotational parameters, $\nu_{\rm s}\sin(i)$ and $\cos(i)$. The prior is especially important for the angle of inclination, which is hardest to infer from the data. We are able to validate the robustness of our asteroseismic results by confirming that their posterior distributions are data-dominated, and not prior-dominated. We can do so by comparing the $68\%$ credible regions of the posterior estimates of $\nu_{\rm s}\sin(i)$ and $i$ against the $68\%$ credible regions of their priors.

A comparison between prior and posterior is shown for 94 stars in $\nu_{\rm s}\sin(i)$, $i$ and $P_{\rm rot}$ in Figure \ref{fig:priors}, arranged by age. In the Figure, results with means and error bars that closely resemble the prior distribution can be interpreted as prior-dominated (i.e. poorly informed by the data). The projected splitting, $\nu_{\rm s}\sin(i)$, is overall well constrained, with only one star appearing prior dominated. This is expected, as the projected splitting is what we observe on the star before decoupling inclination and rotation. The angle of inclination $i$, sampled as $\cos(i)$, more closely follows the prior distribution in most cases. Combining the two, the rotation period $P_{\rm rot}$ has no stars directly corresponding to the effective prior on period, and globally follows a trend with increasing age. The three outliers with fast rotation at late ages (KICs 6603624, 8760414 and 8938364) are discussed in more detail in below.
 blow
The rotation rates as presented in this work are a product of our Bayesian sampling of both projected splitting and angle of inclination. As seen in Figure \ref{fig:priors} there are instances where $i$ or $\nu_{\rm s}\sin(i)$ closely resemble the prior (and are therefore prior-dominated). There are no cases of this when looking at the resulting period measurements, as they will have been informed by at least one strongly data-driven parameter (judging from Figure \ref{fig:priors}, commonly $\nu_{\rm s}\sin(i)$). From this, we conclude that our ensemble of asteroseismic rotation is not strongly dominated by the priors imposed on projected splitting and inclination in our Bayesian analysis.

@@

Conversely, KICs 6603624, 8760414 and 8938364 have extremely slow rotation periods in LEGACY, but extremely \textit{fast} ($<3\, \rm {days}$) in this work. KICs 8760414 and 8938364 are excluded from the gyrochronology analysis below based on checks for $\hat{R}$ and the number of effective samples. Both stars have ages greater than $10\, \rm Gyr$, making their fast rotation rates highly unlikely under any model of rotational evolution. The posterior estimate of $P$ for KIC 6603624 is well-defined in our analysis, but with an age of $7.8\, \rm Gyr$, its measured rotation of $1.2\, \rm days$ is also highly unlikely under any model of rotational evolution. The LEGACY estimate of rotation is similarly extreme at $378\, \rm days$. These three stars have the lowest inclination angles in our sample ($< 10^\circ$), at which point the power in the split components of the seismic modes is so low that it becomes difficult to probe the measure of splitting. The split components for these stars will have a height roughly 3\% of the central mode. For comparison, for the next lowest inclined star at $17^\circ$, this rises to $10\%$. In cases such as these with lower signal-to-noise, spurious peaks may be interpreted as split components.\\

We also compared our asteroseismic estimates of stellar rotation with similar studies in the literature, shown in Figure \ref{fig:literaturecomp}. These included: a study of the binary solar analogues 16 Cyg A \& B \cite{davies+2015}, ; a study of surface and seismic rotation which our catalogue shares 5 stars with \cite{nielsen+2015}, ; and an asteroseismic study of differential rotation with which our catalogue shares 40 targets \cite{benomar+2018}. For the latter, we used their reported splitting value $a_1$, which is equivalent to $\nu_{\rm s}$.

@@

It is of note that 16 Cyg A was found to be rotating slightly faster in our analysis compared to the literature \cite{davies+2015} (deviating within $2\sigma$), despite the fit being performed on the same data. We found an inclination angle that is slightly lower for 16 Cyg A but at a similar projected splitting, which would explain finding a lower value of rotation.

@@

\subsection{Seismic vs spectroscopic rotation}
Different methods of rotation measurement probe different regions of stars. Asteroseismology of main sequence stars probes internal rotation in the near surface layers, where the observed p modes are most sensitive \cite{lund+2014}. Measurements of star-spot modulation instead probe the rotation rates of star spots on the surface. A distinct difference between rotation rates obtained through these different techniques would hold information about differential rotation (both latitudinal and radial) of near-surface layers, such as those we observe in the Sun \cite{beck2000}.

A previous comparison of spectroscopic and seismic rotation rates, performed on a sample of 22 stars, found no significant radial differential rotation. In this work they not only considered rotation rates from spots, but also spectroscopic measures of the projected surface rotation $\textrm{v}\sin(i)$, which they found to be more reliable.

With our expanded sample of asteroseismic rotation we can perform a similar analysis, to both validate our sample and probe radial differential rotation. Figure \ref{fig:vsinilit} shows a comparison between spectroscopic $\textrm{v}\sin(i)$ measurements as listed in LEGACY and Kages (left) and \cite{benomar+2015} (right). In these cases the asteroseismic $\textrm{v}\sin(i)$ has been calculated using our measure of asteroseismic $\nu_s\sin(i)$ and the known asteroseismic radii. Three stars (KICs 6603624, 8760414 and 8938364) have been excluded from this figure due to strong disagreements of measured rotation rates with the literature (see above).

For the LEGACY and Kages sample, we find no strong deviation from the 1:1 line except at very low velocities, which is likely due to biases inherent to spectroscopic line broadening measurements \cite{doyle+2014, tayar+2015}. For the \cite{benomar+2015} sample stars lie a lot closer to the 1:1 line.

Overall, there appears to be a global offset where spectroscopic measurements of projected rotation appear faster than asteroseismic measures. Based on the LEGACY and Kages $v\sin(i)$ values, this offset is roughly $18\%$ (i.e. spectroscopic projected rotation rates are higher than asteroseismic rates). This offset is much smaller ($\sim5\%$) for the \cite{benomar+2015} sample, albeit for far fewer stars. These offsets are within the typical disagreement between spectroscopic methods, based on comparisons of projected rotation measurements for red giant stars \cite[see Figure 2]{tayar+2015}, especially at $< 5\, \rm{km\,s^{-1}}$.

It is worth addressing the impact this comparison would have on our conclusions for gyrochronology, were we to take the spectroscopic rotation rates as if they were the truth, even at velocities of $< 5\, \rm{km\,s^{-1}}$. For the majority of cases, the seismic velocities are slower than the spectroscopic velocities, which would bias our ensemble towards favouring a standard evolution over a weakened magnetic braking scenario (as stars would overall appear to be rotating slower at late ages). Based on the comparison to spot rotation presented in the main body of this paper, and given known issues comparing seismic and spectroscopic rotation at low velocities, we conclude that the divergence seen in Figure \ref{fig:vsinilit} does not undermine our results.

@@

Do this

\section{Verifying consequences for gyrochronology}
\subsection{Limits of our stellar models}\label{ssec:limits}
The models presented in \cite{vansaders+2019} are constructed for metallicities of $-0.4 < \rm{[Fe/H]} < 0.4\, \rm{dex}$, in steps of $0.1\, \rm {dex}$. Our sample of 91 stars contained 4 stars with metallicities below $-0.4\, \rm{dex}$, which are shown as shaded symbols in Figure \ref{fig:fullmodels}. Of these 4, 3 were included in our final sample of 73 stars used to evaluate our stellar models: KICs 7970740, 8684723 and KIC 9965715. All three are classed as MS stars, with metallicities of $-0.54 \pm 0.10$, $-0.42 \pm 0.10$ and $-0.44 \pm 0.18\, \rm {dex}$ respectively, placing them within $3\sigma$ of the metallicity limits of our stellar models. KICs 8684723 and 9965715 strongly agree with the WMB model, whereas KIC 7970740 weakly prefers the standard model. Excluding these stars was not found to significantly alter the joint posterior distribution shown in Figure \ref{fig:modelresults}.
